A data-driven, pixel-level model to improve the precision of \Kepler\ photometry

### 2013-10-17 documents/plm.tex
start terminology cray cray

\newcommand{\notenglish}[1]{\textit{#1}}
\newcommand{\sic}{\notenglish{sic}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\Kepler}{\project{Kepler}}
\newcommand{\name}{PLM}
\newcommand{\todo}[1]{\textbf{#1}}

@@

Before we start, a few reminders about the \Kepler\ data are in order:
The satellite always observes precisely the same field, at fixed pointing (as closely as possible).
The satellite is rolled by 90~deg every 90-ish days (to satisfy Solar-Angle constraints).
The PSF varies strongly across the field and is badly sampled.
The stars span a huge range in brightness;
  some of \Kepler's most important targets even saturate the device and bleed charge.
The stellar photometry returned by the \Kepler\ SAP and PDC pipelines is based on
  straight, \emph{unweighted} sums of pixels in small patches centered on the stellar centroid.
We will return to this latter point at the end;
  this kind of photometry cannot be optimal;
  it must be possible to do a better job with the photometric measurements.
That's beyond the scope of this paper but a place for a valuable intervention on the \Kepler\ data.

@@

In this framework, the data used to \emph{train} the model%
  ---meaning set the values of the parameters of the model---%
  are disjoint from the \emph{test} data
  ---meaning the data that are being predicted (the target data).
Because we are extremely concerned with detecting Earth-analog exoplanet transits,
  which take about ten hours,
  we adopt an extreme version of the train-and-test framework,
  in which the training data are always separated from the test data by at least ten hours.
That is, when we are using the model to predict a particular pixel in the target data taken at time $t$,
  we use parameters obtained by an optimization that makes use of only data
  that comes either at times prior to $t-\Delta t$ or after $t+\Delta t$,
  where $\Delta t$ is a tunable parameter but which we will set to roughly $10$~h.
This ensures that no information about any exoplanet transit itself
  can be entering into the prediction of the pixels contributing to the stellar photometry.
In general, if there is a scientific goal of preserving intrinsic stellar variability,
  or transit signals,
  on time-scales of $\tau$,
  the parameter optimization ought to be based on training data taken with a time-exclusion zone of half-width $\Delta t > \tau$.

@@

For each pixel, we choose the set of basis-vector pixels how?

For each time-step, we choose the set of training time steps how?

>>>>

\newcommand{\notenglish}[1]{\textit{#1}}
\newcommand{\sic}{\notenglish{sic}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\Kepler}{\project{Kepler}}
\newcommand{\name}{PLM}
\newcommand{\set}[1]{\mathcal{#1}}
\newcommand{\todo}[1]{\textbf{#1}}

@@

Before we start, a few reminders about the \Kepler\ data are in order:
The satellite always observes precisely the same field, at fixed pointing (as closely as possible).
Only a few percent of the 80 million pixels in the focal plane are telemetered down
  to Earth from each 30-min exposure;
  the telemetered pixels are associated with \Kepler\ target stars chosen for study by the \Kepler\ Team.
The satellite is rolled by 90~deg every 90-ish days (to satisfy Solar-Angle constraints).
The focal plane contains many CCDs;
  this plus the 90-deg rotations means that each star is at a particular location
  on a particular CCD for 90-ish-day contiguous periods of time.
The PSF varies strongly across the field and is badly sampled.
The stars span a huge range in brightness;
  some of \Kepler's most important targets even saturate the device and bleed charge.
The stellar photometry returned by the \Kepler\ SAP and PDC pipelines is based on
  straight, \emph{unweighted} sums of pixels in small patches centered on the stellar centroid.
We will return to this latter point at the end;
  this kind of photometry cannot be optimal;
  it must be possible to do a better job with the photometric measurements.
That's beyond the scope of this paper but a place for a valuable intervention on the \Kepler\ data.

@@

In this framework, the data used to \emph{train} the model%
  ---meaning set the values of the parameters of the model---%
  are disjoint from the \emph{test} data
  ---meaning the data that are being predicted (the target data).
Because we are extremely concerned with detecting Earth-analog exoplanet transits,
  which take about ten hours,
  we adopt an extreme version of the train-and-test framework,
  in which the training data are always separated from the test data by at least ten hours.
That is, when we are using the model to predict a particular pixel in the target data taken at time $t$,
  we use parameters obtained by an optimization that makes use of only data
  that comes either at times prior to $t-\Delta t$ or after $t+\Delta t$,
  where $\Delta t$ is a tunable parameter but which we will set to roughly $12$~h.
This ensures that no information about any exoplanet transit itself
  can be entering into the prediction of the pixels contributing to the stellar photometry.
In general, if there is a scientific goal of preserving intrinsic stellar variability,
  or transit signals,
  on time-scales of $\tau$,
  the parameter optimization ought to be based on training data taken with a time-exclusion zone of half-width $\Delta t > \tau$.

@@

Each individual \Kepler\ target star is, for each quarter (each 90-ish-day period),
  on a particular CCD on the \Kepler\ focal plane.
Associated with each target star is some set of pixels,
  located in a small contiguous patch more-or-less centered on the target star,
  and telemetered down once for every 30-min exposure.
(Here we are ignoring all short-cadence data that is taken in a different mode with shorter exposures.)
That is, each telemetered pixel in each CCD is associated with a particular \Kepler\ target star.

Let's consider a particular pixel $m$ in the focal plane in one particular quarter.
In that quarter, the spacecraft delivers $N\sim 4000$ measurements $I_{mn}$
  of the intensity falling in that pixel at the $N$ times $t_n$ at which exposures were taken during the quarter.
(For intensity measurements $I_{mn}$ here we are using pixels from the \todo{WHICH \Kepler\ image product?})
We want to build a prediction for the intensity $I_{mn}$ in pixel $m$ at each time $t_n$
  using the intensities $I_{m'n}$ of other pixels $m'$.
The question is:  What other pixels to choose?
There are many possible qualitatively and quantitatively different choices here.
Somewhat arbitrarily,
  we choose all the pixels $m'$ associated with the $Q=5$ stars on the same CCD in the same quarter
  that are closest in magnitude
  (\Kepler\ magnitude as reported in the \Kepler\ Input Catalog)
  to the target star associated with pixel $m$.
We call this set of pixels%
  ---from the $Q$ stars on the same CCD---%
  the \emph{pixel set} $\set{M}_m$ associated with pixel $m$.
Note that because the set $\set{M}_m$ is of pixels associated with \emph{different} stars
  than the star associated with pixel $m$,
  the pixel $m$ will not be in the set $\set{M}_m$,
  and nor will any of pixel $m$'s close neighbors.
That is, there will be no (or almost no) overlap in stellar illumination of pixel $m$
  and the pixels $m'\in\set{M}_m$.

When we are predicting the measurement $I_{mn}$ from pixel $m$ at a particular time $t_n$,
  we are going to use a train-and-test framework in which not only do we not use
  data at time $t_n$ in optimizing the parameters of the regression or fit,
  but we don't even use any times $t$ such that $|t-t_n| < \Delta t$,
  where $\Delta t\sim 12$\,h.
That is, we train (optimize fit parameters) using a set $\set{N}_n$ of time
  indices $n'$ such that for all $n'\in\set{N}_n$,
  $t_{n'}$ is in the same quarter as $t_n$,
  and $|t_{n'} - t_n|>\Delta t$.
The training set of time indices $\set{N}_n$ therefore does not overlap the time point $t_n$,
  nor any of its neighbors in a time window of half-width $\Delta_t$.

\begin{eqnarray}
I_{mn}         &=& I^{\ast}_{mn} + \mbox{noise}
\\
I^{\ast}_{mn}  &=& \sum_{m'\in\set{M}_m} a_{mm'}\,I_{m'n}
\\
I^{\ast}_{mn'} &=& \sum_{m'\in\set{M}_m} a_{mm'}\,I_{m'n'}
\\
\chi^2_{mn}    &=& \sum_{n'\in\set{N}_n} \frac{[I_{mn'} - I^{\ast}_{mn'}]^2}{\sigma^2_{mn'}}
                 + \sum_{m'\in\set{M}_m} \frac{a_{mm'}^2}{\Sigma^2}
\end{eqnarray}
